A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions

### 2022-01-10 2022/OM&S/paper.tex
Corrections in Section 1

For optimization of convolutions a wide spectrum of algorithms can be used in dependence on convolution properties generated by the features of criteria in the initial MCO problem. In particular, for the multiextremal case where it is required to find the global optimum the publications \cite{Evtushenko2014,Zilinskas2015,GERGEL2017_1,Gergel2019_2,Barkalov2021} describe applications of the global optimization algorithms to MCO problems in combination with the convolution approach.

Among other methods of solving MCO problems it is necessary to pay attention to the important class of metaheuristic nature-inspired algorithms. Many of them are nature-inspired, i.e.,  are  based on modeling physical processes (for example, simulated annealing \cite{Locatelli2002,Aarts2014}) or mainly behavior of biological agents including evolutionary \cite{Price2005,Coello2007} and as their part genetic algorithms \cite{Ruiz2015,Deb2002,Zitzler2001} and particle swarm methods modeling the swarming intelligence of bees, fireflies, birds, ants, etc. \cite{Mostaghim2007,Nebro2009,Durillo2010}.

@@

\textcolor[rgb]{1,0,0}{If number of completed trials is not evenly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then }
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
\end{equation}
Here
\begin{gather}\label{eq:18}
B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q-1} + v_q}{m},\\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{gather}
is minimal computed value of the objective function $\varphi(t)$ at trial points,
\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}
is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}
is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

@@

Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ being scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).

>>>>

For optimization of convolutions a wide spectrum of algorithms can be used in dependence on convolution properties generated by the features of criteria in the initial MCO problem. In particular, for the multiextremal case where it is required to find the global optimum the publications \cite{Evtushenko2014,Zilinskas2015,GERGEL2017_1,Gergel2019_2,Barkalov2021} describe applications of the global optimization algorithms to MCO problems in combination with the convolution approach. Among these methods of multiextremal optimization the efficient algorithms that utilize  schemes of dimensionality reduction such as Peano mappings \cite{Strongin2000,Sergeyev2013} or nested optimization \cite{Strongin2000,Grishagin2018} are perspective for application in the  frameworks of multiobjective optimization.

As for other methods of solving MCO problems it is necessary to pay attention to the important class of metaheuristic nature-inspired algorithms. Many of them are nature-inspired, i.e.,  are  based on modeling physical processes (for example, simulated annealing \cite{Locatelli2002,Aarts2014}) or mainly behavior of biological agents including evolutionary \cite{Price2005,Coello2007} and as their part genetic algorithms \cite{Ruiz2015,Deb2002,Zitzler2001} and particle swarm methods modeling the swarming intelligence of bees, fireflies, birds, ants, etc. \cite{Mostaghim2007,Nebro2009,Durillo2010}.

@@

If number of completed trials is not exactly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
\end{equation}
Here
\begin{gather}\label{eq:18}
B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q-1} + v_q}{m},\\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{gather}
is minimal computed value of the objective function $\varphi(t)$ at trial points,
\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}
is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}
is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

@@

Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ which are scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}). This procedure will be called Repeated Algorithm with Local Refinements and Data Accumulation (RALR-DA).
